\bf Appendix: Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-01-17 appendix.qmd
Updated Figure 4

---
title: "Appendix: Looking at Non-Linear Dimension Reductions as Models in the Data Space"
format: 
    jasa-pdf:
        keep-tex: true
    jasa-html: default
author:
  - name: Jayani P.G. Lakshika
    affiliations:
      - name: Monash University
        department: Econometrics & Business Statistics
        address: Clayton
        city: VIC 
        country: Australia
        postal-code: 3800
    orcid: 0000-0002-6265-6481
    email: jayani.piyadigamage@monash.edu
    url: https://jayanilakshika.netlify.app/
  - name: Dianne Cook
    affiliations:
      - name: Monash University
        department: Econometrics & Business Statistics
        address: Clayton
        city: VIC 
        country: Australia
        postal-code: 3800
    orcid: 0000-0002-3813-7155
    email: dicook@monash.edu 
    url: http://www.dicook.org/
  - name: Paul Harrison
    affiliations:
      - name: Monash University
        department: MGBP, BDInstitute
        address: Clayton
        city: VIC 
        country: Australia
        postal-code: 3800
    orcid: 0000-0002-3980-268X
    email: 	paul.harrison@monash.edu
    url: 
  - name: Michael Lydeamore
    affiliations:
      - name: Monash University
        department: Econometrics & Business Statistics
        address: Clayton
        city: VIC 
        country: Australia
        postal-code: 3800
    orcid: 0000-0001-6515-827X
    email: michael.lydeamore@monash.edu
    url: https://www.michaellydeamore.com/
  - name: Thiyanga S. Talagala
    affiliations:
      - name: University of Sri Jayewardenepura
        department: Statistics
        address: Gangodawila
        city: Nugegoda 
        country: Sri Lanka
        postal-code: 10100
    orcid: 0000-0002-0656-9789
    email: ttalagala@sjp.ac.lk 
    url: https://thiyanga.netlify.app/
tbl-cap-location: bottom
header-includes: | 
  \usepackage{amsmath}
  \usepackage{float}
  \usepackage{hyperref}
  \usepackage[utf8]{inputenc}
  \usepackage{bm}
  \def\tightlist{}
  \usepackage{setspace}
  \newcommand\pD{$p\text{-}D$}
  \newcommand\kD{$k\text{-}D$}
  \newcommand\dD{$d\text{-}D$}
  \newcommand\gD{$2\text{-}D$}
  \newcommand{\To}{\textbf{to}~}
  \usepackage{abstract}
  \renewcommand{\abstractname}{}    % clear the title
  \renewcommand{\absnamepos}{empty} % originally center
---

@@

```{r}
#| label: fig-pbmc-mse-umap
#| fig-pos: H
#| out-width: 100%
#| fig-cap: "Assessing which of the 2 UMAP layouts with different hyperparameter settings (n_neighbors: $30$, min_dist: $0.3$ (red), n_neighbors: $80$, min_dist: $0.5$ (blue)) on the PBMC data is the better representation using MSE for varying binwidth ($a_1$). Colour used for the lines and points in the left plot and in the scatterplots represents UMAP layout (a, b). Of the two, layout b is optimal across all binwidths making it the best choice."

@@

```{r}
#| echo: false
#| eval: false
#| fig-cap: "Model in \\gD{}, on the layout a, and two views of the fit in projections from $50\\text{-}D$, for the PBMC data ($n =  5858$ and $p = 50$). Layout shows close clusters. In $50\\text{-}D$, the data shows "
#| label: fig-model-pbmc-author-proj-scdeed
#| fig-pos: H
#| fig-width: 15
#| fig-height: 5
#| out-width: 80%

>>>>

---
title: "Appendix: Looking at Non-Linear Dimension Reductions as Models in the Data Space"
format: 
    jasa-pdf:
        keep-tex: true
    jasa-html: default
author:
  - name: Jayani P.G. Lakshika
    affiliations:
      - name: Monash University
        department: Econometrics & Business Statistics
        address: Clayton
        city: VIC 
        country: Australia
        postal-code: 3800
    orcid: 0000-0002-6265-6481
    email: jayani.piyadigamage@monash.edu
    url: https://jayanilakshika.netlify.app/
  - name: Dianne Cook
    affiliations:
      - name: Monash University
        department: Econometrics & Business Statistics
        address: Clayton
        city: VIC 
        country: Australia
        postal-code: 3800
    orcid: 0000-0002-3813-7155
    email: dicook@monash.edu 
    url: http://www.dicook.org/
  - name: Paul Harrison
    affiliations:
      - name: Monash University
        department: MGBP, BDInstitute
        address: Clayton
        city: VIC 
        country: Australia
        postal-code: 3800
    orcid: 0000-0002-3980-268X
    email: 	paul.harrison@monash.edu
    url: 
  - name: Michael Lydeamore
    affiliations:
      - name: Monash University
        department: Econometrics & Business Statistics
        address: Clayton
        city: VIC 
        country: Australia
        postal-code: 3800
    orcid: 0000-0001-6515-827X
    email: michael.lydeamore@monash.edu
    url: https://www.michaellydeamore.com/
  - name: Thiyanga S. Talagala
    affiliations:
      - name: University of Sri Jayewardenepura
        department: Statistics
        address: Gangodawila
        city: Nugegoda 
        country: Sri Lanka
        postal-code: 10100
    orcid: 0000-0002-0656-9789
    email: ttalagala@sjp.ac.lk 
    url: https://thiyanga.netlify.app/
tbl-cap-location: bottom
bibliography: bibliography.bib 
header-includes: | 
  \usepackage{amsmath}
  \usepackage{float}
  \usepackage{hyperref}
  \usepackage[utf8]{inputenc}
  \usepackage{bm}
  \def\tightlist{}
  \usepackage{setspace}
  \newcommand\pD{$p\text{-}D$}
  \newcommand\kD{$k\text{-}D$}
  \newcommand\dD{$d\text{-}D$}
  \newcommand\gD{$2\text{-}D$}
  \newcommand{\To}{\textbf{to}~}
  \usepackage{abstract}
  \renewcommand{\abstractname}{}    % clear the title
  \renewcommand{\absnamepos}{empty} % originally center
---

@@

```{r}
#| label: fig-pbmc-mse-umap
#| fig-pos: H
#| out-height: 50%
#| fig-cap: "Assessing which of the 2 UMAP layouts with different hyperparameter settings (n_neighbors: $30$, min_dist: $0.3$ (red), n_neighbors: $80$, min_dist: $0.5$ (blue)) on the PBMC data is the better representation using MSE for varying binwidth ($a_1$). Colour used for the lines and points in the left plot and in the scatterplots represents UMAP layout (a, b). Of the two, layout b is optimal across all binwidths making it the best choice."

@@

```{r}
#| echo: false
#| fig-cap: "Model in \\gD{}, on the layout a, and two views of the fit in projections from $50\\text{-}D$, for the PBMC data ($n =  5858$ and $p = 50$). Layout shows close clusters. In $50\\text{-}D$, the data shows "
#| label: fig-model-pbmc-author-proj-scdeed
#| fig-pos: H
#| fig-width: 15
#| fig-height: 5
#| out-width: 30%

### 2025-01-17 appendix.tex
Updated Figure 4

\textbf{?@fig-model-pbmc-author-proj-scdeed} shows layout from the
original published article along with that suggested by scDEED to be
more accurate. The log MSE vs binwidth plot illustrates that our
approach would suggest that scDEED is correct here, that layout b is
more accurately reflecting the cluster structure in the data.

@@

\includegraphics[width=1\textwidth,height=\textheight]{appendix_files/figure-pdf/fig-pbmc-mse-umap-1.pdf}

@@

>>>>

Figure~\ref{fig-model-pbmc-author-proj-scdeed} shows layout from the
original published article along with that suggested by scDEED to be
more accurate. The log MSE vs binwidth plot illustrates that our
approach would suggest that scDEED is correct here, that layout b is
more accurately reflecting the cluster structure in the data.

@@

\includegraphics[width=1\textwidth,height=0.5\textheight]{appendix_files/figure-pdf/fig-pbmc-mse-umap-1.pdf}

@@

\includegraphics[width=0.3\textwidth,height=\textheight]{appendix_files/figure-pdf/fig-model-pbmc-author-proj-scdeed-1.pdf}

\caption{\label{fig-model-pbmc-author-proj-scdeed}Model in \gD{}, on the
layout a, and two views of the fit in projections from \(50\text{-}D\),
for the PBMC data (\(n =  5858\) and \(p = 50\)). Layout shows close
clusters. In \(50\text{-}D\), the data shows}

  \bibliography{bibliography.bib}
